Fake News vs Fox News

### 2024-04-19 paper/Fake_News_vs_Fox_News.qmd
final changes, small grammar edits

# Introduction
The 2020 U.S. presidential election, between Republican President Donald Trump, and Democratic candidate Joe Biden, was one of the most intense in recent history. This was especially due to the large polarization in political opinion among Americans. This was especially reflected in the media outlets favored by Americans, with Fox News being commonly known as the news network highly favored by Republicans, and Cable News Network (CNN) being commonly known as the Democrats' favorite.

Throughout his presidency and re-election campaign, Trump held noticeable animosity against liberal networks. Trump famously labelled any unflattering issues that might work against him as "Fake News" [@timm2020trump]. He even handed out the "Fake News Award" through Twitter to several notable liberal media outlets, including CNN [@siddiqui2018trump]. Democrats also have long-standing issues with the Republican news networks like Fox News. For instance, Joe Biden has referred to Fox News as "one of the most destructive forces in the United States" [@stelter2022biden]. With certain networks having clear political alliances, these different media outlets could influence their viewers' political ideologies and voting decisions. Consequently, it is important to explore whether US media networks have had significant influences on election outcomes. Understanding this could give a  good understanding of this relationship could enable voters to be more critical of the media they consume.

@@

The estimand of this paper is the probability of people voting Biden during the 2020 election based on news channel viewership. However, due to the difficulty of collecting data on millions of voters, this paper estimates the estimand using the logistic regression model trained using a sample from the CES dataset.

The findings indicate that watching media networks with specific political ideologies is associated with higher chances of voting for the candidate they prefer. This is most apparent in with CNN and Fox News, who are shown to exert heavy influence on voters identifying as "Independent" or "Other.". Watching local news over these national networks appears to have no significant effect on the probability of voting for Trump or Biden, as well.

@@

The dependent variable is: `presvote20post`, renamed to `voted_for`. This variable represents the presidential candidate the respondent voted for in the 2020 election in the form of a numerical variable. The variable `votereg` represents whether the respondent is registered to vote using numerical numbers. This paper will only analyze respondents who are registered to vote and focus on the outcome of two candidates, Joe Biden representing the Democratic party and Donald Trump representing the Republican party. To analyze the observations of interest, I first limited the observations to the ones that responded "Yes" in `votereg`. And I filtered the responses in `presvote20post` to only "Biden" or "Trump" and converted the variable into a dummy variable where 1 represents "Biden" and 0 represents "Trump". The rest of the variables this paper focuses on are being divided into three categories: Media Use - Networks, Media Use - TV Type, and Party Affiliation.

@@

### Model justification
A positive relationship is expected between watching left-leaning (Democrat) media networks and voting for Biden, as well as a similar positive relationship between watching right-leaning (Republican) media networks and voting for Trump. According to @grieco2020americans, among the 7 networks identified in this study, Fox News is known as the most conservative media network, while CNN and MSNBC are leading liberal media outlets. Slightly leaning left, ABC, CBS, and NBC are generally considered neutral. PBS is also more liberal and Democratic-leaning, but not as extreme as CNN or MSNBC [@blake2014ranking]. This assumption is grounded in findings that suggest exposure to media networks influences voters [@wettstein2017media].

>>>>

# Introduction
The 2020 U.S. presidential election, between Republican President Donald Trump and Democratic candidate Joe Biden, was one of the most intense in recent history. A key factor behind this can be attributed to the large polarization in political opinion among Americans. This was especially reflected in the media outlets favored by Americans, with Fox News being commonly known as the news network highly favored by Republicans, and Cable News Network (CNN) being commonly known as the Democrats' favorite.

Throughout his presidency and re-election campaign, Trump held noticeable animosity against liberal networks. Trump famously labelled any unflattering issues that might work against him as "Fake News" [@timm2020trump]. He even handed out the "Fake News Award" through Twitter to several notable liberal media outlets, including CNN [@siddiqui2018trump]. Democrats also have long-standing issues with Republican news networks like Fox News. For instance, Joe Biden has referred to Fox News as "one of the most destructive forces in the United States" [@stelter2022biden]. With certain networks having clear political alliances, these different media outlets could influence their viewers' political ideologies and voting decisions. Consequently, it is important to explore whether US media networks have had significant influences on election outcomes. A good understanding of this relationship could enable voters to be more critical of the media they consume.

@@

The estimand of this paper is the probability of people voting for Biden during the 2020 election based on news channel viewership. However, due to the difficulty of collecting data on millions of voters, this paper estimates the estimand using the logistic regression model trained using a sample from the CES dataset.

The findings indicate that watching media networks with specific political ideologies is associated with higher chances of voting for the candidate they prefer. This is most apparent in with CNN and Fox News, who are shown to exert heavy influence on voters identifying as "Independent" or "Other." Additionally, this paper also studies the influence of watching national newscasts on voting outcomes compared to watching both national and local newscasts, finding no significant differences between the outcomes of the two.

@@

The dependent variable is: `presvote20post`, renamed to `voted_for`. This variable represents the presidential candidate the respondent voted for in the 2020 election in the form of a numerical variable. The variable `votereg` represents whether the respondent is registered to vote using numerical numbers. This paper will only analyze respondents who are registered to vote and focus on the outcome of two candidates, Joe Biden and Donald Trump. To analyze the observations of interest, I first limited the observations to the ones that responded "Yes" in `votereg`. And I filtered the responses in `presvote20post` to only "Biden" or "Trump" and converted the variable into a dummy variable where 1 represents "Biden" and 0 represents "Trump". The rest of the variables this paper focuses on are being divided into three categories: Media Use - Networks, Media Use - TV Type, and Party Affiliation.

@@

## Model justification
A positive relationship is expected between watching left-leaning (Democrat) media networks and voting for Biden, as well as a similar positive relationship between watching right-leaning (Republican) media networks and voting for Trump. According to @grieco2020americans, among the 7 networks identified in this study, Fox News is known as the most conservative media network, while CNN and MSNBC are leading liberal media outlets. Slightly leaning left, ABC, CBS, and NBC are generally considered neutral. PBS is also more liberal and Democratic-leaning, but not as extreme as CNN or MSNBC [@blake2014ranking]. This assumption is grounded in findings that suggest exposure to media networks influences voters [@wettstein2017media].
